DATA 624 Homework 9

### 2020-04-21 DATA624/HW9.Rmd
More work on HW9

```{r setup, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA,
                      message = FALSE, warning = FALSE)
library(tidyverse)
library(kableExtra)
library(tidyr)
library(RColorBrewer)
library(gbm)
library(randomForest)
library(caret)
library(party)
library(rpart)
library(mlbench)
knitr::knit_hooks$set(inline = function(x) {
  if (!is.numeric(x)) {
    x
  } else {
    prettyNum(round(x, 2), big.mark = ",")
  }
})
options(scipen = 999)

@@

pls_model <- train(
  Yield ~ ., data = train_df, method = "pls",
  center = TRUE,
  scale = TRUE,
  trControl = trainControl("cv", number = 10),
  tuneLength = 25
)

pls_predictions <- predict(pls_model, test_df)

pls_in_sample <- pls_model$results[pls_model$results$ncomp == pls_model$bestTune$ncomp,]
results <- data.frame(t(postResample(pred = pls_predictions, obs = test_df$Yield))) %>%
  mutate("In Sample RMSE" = pls_in_sample$RMSE,
         "In Sample Rsquared" = pls_in_sample$Rsquared,
         "In Sample MAE" = pls_in_sample$MAE,
         "Model"= "PLS")
```

**I'm going to fit the data against a series of tree based models.  I will use caret with vanilla settings to see how well the preform against a tuned PLS model.**

```{r}
gbm_model <- train(Yield ~ ., data = train_df, method="gbm", verbose = FALSE)
gbm_predictions <- predict(gbm_model, test_df)
gbm_in_sample <- merge(gbm_model$results, gbm_model$bestTune)
results <- data.frame(t(postResample(pred = gbm_predictions, obs = test_df$Yield))) %>%
  mutate("In Sample RMSE" = gbm_in_sample$RMSE,
         "In Sample Rsquared" = gbm_in_sample$Rsquared,
         "In Sample MAE" = gbm_in_sample$MAE,
         "Model"= "Boosted Tree") %>%
  rbind(results)
```

@@

```{r}
rf_model <- train(Yield ~ ., data = train_df, method = "rf")
rf_predictions <- predict(rf_model, test_df)
rf_in_sample <- merge(rf_model$results, rf_model$bestTune)
results <- data.frame(t(postResample(pred = rf_predictions, obs = test_df$Yield))) %>%
  mutate("In Sample RMSE" = rf_in_sample$RMSE,
         "In Sample Rsquared" = rf_in_sample$Rsquared,
         "In Sample MAE" = rf_in_sample$MAE,
         "Model"= "Random Forest") %>%
  rbind(results)
```

#### Cubist Model

```{r}
cubist_model <- train(Yield ~ ., data = train_df, method="cubist")

cubist_predictions <- predict(cubist_model, test_df)
cubist_in_sample <- merge(cubist_model$results, cubist_model$bestTune)
results <- data.frame(t(postResample(pred = cubist_predictions, obs = test_df$Yield))) %>%
  mutate("In Sample RMSE" = cubist_in_sample$RMSE,
         "In Sample Rsquared" = cubist_in_sample$Rsquared,
         "In Sample MAE" = cubist_in_sample$MAE,
         "Model"= "Cubist") %>%
  rbind(results)
```

@@

```{r}
results %>%
  kable() %>%
  kable_styling()
```

**All tree model's fit the in sample data better than the PLS model.  However they may have overfit the data as the RMSE on the test set is higher for the random forest and boosted tree.  The cubist model is the only on that out preformed the PLS on the test set.  These findings are on vanilla models.  Additional tuning could be used to improve the model's performance**

@@

```{r}
varImp(cubist_model, 10)
```

**The manufacturing process variables continue to dominate the list.  Some of the variables on on both lists (in different orders of importance) while there are differences between the two lists:**

```{r}
varImp(pls_model, 10)
```

@@

>>>>

```{r setup, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA,
                      message = FALSE, warning = FALSE)
library(tidyverse)
library(kableExtra)
library(tidyr)
library(RColorBrewer)
library(gbm)
library(randomForest)
library(caret)
library(rattle)
library(party)
library(rpart)
library(mlbench)
knitr::knit_hooks$set(inline = function(x) {
  if (!is.numeric(x)) {
    x
  } else {
    prettyNum(round(x, 2), big.mark = ",")
  }
})
options(scipen = 999)

@@

pls_model <- train(
  Yield ~ ., data = train_df, method = "pls",
  center = TRUE,
  scale = TRUE,
  trControl = trainControl("cv", number = 10),
  tuneLength = 25
)
pls_predictions <- predict(pls_model, test_df)
pls_in_sample <- pls_model$results[pls_model$results$ncomp == pls_model$bestTune$ncomp,]
results <- data.frame(t(postResample(pred = pls_predictions, obs = test_df$Yield))) %>%
  mutate("In Sample RMSE" = pls_in_sample$RMSE,
         "In Sample Rsquared" = pls_in_sample$Rsquared,
         "In Sample MAE" = pls_in_sample$MAE,
         "Model"= "PLS")
pls_model
```

**I'm going to fit the data against a series of tree based models.  I will use caret and try to match the same paramaters to guague th preformance against the similar PLS model.**

#### Bagged Tree

```{r}
set.seed(42)
bagControl = bagControl(fit = ctreeBag$fit, predict = ctreeBag$pred, aggregate = ctreeBag$aggregate)
bag_model <- train(Yield ~ ., data = train_df, method="bag", bagControl = bagControl,
                   center = TRUE,
                   scale = TRUE,
                   trControl = trainControl("cv", number = 10),
                   tuneLength = 25)
bag_predictions <- predict(bag_model, test_df)
bag_in_sample <- merge(bag_model$results, bag_model$bestTune)
results <- data.frame(t(postResample(pred = bag_predictions, obs = test_df$Yield))) %>%
  mutate("In Sample RMSE" = bag_in_sample$RMSE,
         "In Sample Rsquared" = bag_in_sample$Rsquared,
         "In Sample MAE" = bag_in_sample$MAE,
         "Model"= "Bagged Tree") %>%
  rbind(results)
bag_model
```
#### Boosted Tree

```{r}
set.seed(42)
gbm_model <- train(Yield ~ ., data = train_df, method="gbm", verbose = FALSE,
                   trControl = trainControl("cv", number = 10),
                   tuneLength = 25)
gbm_predictions <- predict(gbm_model, test_df)
gbm_in_sample <- merge(gbm_model$results, gbm_model$bestTune)
results <- data.frame(t(postResample(pred = gbm_predictions, obs = test_df$Yield))) %>%
  mutate("In Sample RMSE" = gbm_in_sample$RMSE,
         "In Sample Rsquared" = gbm_in_sample$Rsquared,
         "In Sample MAE" = gbm_in_sample$MAE,
         "Model"= "Boosted Tree") %>%
  rbind(results)
gbm_model
```

@@

```{r}
set.seed(42)
rf_model <- train(Yield ~ ., data = train_df, method = "ranger", 
                  scale = TRUE,
                  trControl = trainControl("cv", number = 10),
                  tuneLength = 25)
rf_predictions <- predict(rf_model, test_df)
rf_in_sample <- merge(rf_model$results, rf_model$bestTune)
results <- data.frame(t(postResample(pred = rf_predictions, obs = test_df$Yield))) %>%
  mutate("In Sample RMSE" = rf_in_sample$RMSE,
         "In Sample Rsquared" = rf_in_sample$Rsquared,
         "In Sample MAE" = rf_in_sample$MAE,
         "Model"= "Random Forest") %>%
  rbind(results)
rf_model
```

#### Conditional Inference Random Forest

```{r}
set.seed(42)
crf_model <- train(Yield ~ ., data = train_df, method = "cforest",
                   trControl = trainControl("cv", number = 10),
                   tuneLength = 25)
crf_predictions <- predict(crf_model, test_df)
crf_in_sample <- merge(crf_model$results, crf_model$bestTune)
results <- data.frame(t(postResample(pred = crf_predictions, obs = test_df$Yield))) %>%
  mutate("In Sample RMSE" = crf_in_sample$RMSE,
         "In Sample Rsquared" = crf_in_sample$Rsquared,
         "In Sample MAE" = crf_in_sample$MAE,
         "Model"= "Conditional Random Forest") %>%
  rbind(results)
crf_model
```

@@

```{r}
results %>%
  arrange(RMSE) %>%
  kable() %>%
  kable_styling()
```

**All tree model's fit the in sample data better than the PLS model.  However they may have overfit the data as the RMSE on the test set is higher for the random forest and boosted tree.  The boosted tree model is the only on that out preformed the PLS on the test set.**

@@

```{r}
varImp(bag_model)
```

**The manufacturing process variables continue to dominate the list.  The same variables are found in both lists (in different orders of importance)**

```{r}
varImp(pls_model)
```

@@

```{r}
set.seed(1)
cart_model <- train(Yield ~ ., data = train_df, method = "rpart",
                    trControl = trainControl("cv", number = 10),
                    tuneLength = 25)
fancyRpartPlot(cart_model$finalModel, sub="")
```

**This indicates that if you want to maximize yield, then the manufacturing process 32 must be greater than or equal to 0.19, process 13 < -0.85 and the Biological material 3 >= 0.49.  This recipie will produce the greatest yield.  Caution should be used however as this model is likely overfitting the training data.**
